---
title: "aquamaps_groups.Rmd"
author: "Ben Best"
date: "11/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

original: `aquamaps_ingest.Rmd`

## Taxonomic Groups

Tittensor DP, C Mora, W Jetz, HK Lotze, D Ricard, EV Berghe, B Worm (2010) [Global patterns and predictors of marine biodiversity across taxa](https://doi.org/10.1038/nature09329). _Nature_

- Primarily coastal species (Tittensor et al 2010)
    - **Coastal fishes**: `Class == "Actinopterygii" & !Family %in% c("Scombridae", "Istiophoridae", "Xiphiidae")`
    - Non-oceanic sharks: TODO with `rredlist::rl_habitats()`
    - **Non-squid cephalopods**: `Class == "Cephalopoda" & !Order %in% c("Myopsida","Oegopsida")`
    - **Pinnipeds**: `Family %in% c("Odobenidae", "Otariidae", "Phocidae")`
    - **Corals**: `Class == "Anthozoa"` (soft/hard + anemones + gorgonians)
    - **Seagrasses**: `Order == "Alismatales"`
    - **Mangroves**: `Genus == "Avicennia" & Species == "germinans"` (n=1)
- Primarily oceanic species (Tittensor et al 2010)
    - **Tunas and billfishes**: `Family %in% c("Scombridae", "Istiophoridae", "Xiphiidae")`
    - Oceanic sharks: TODO with `rredlist::rl_habitats()`
    - **Squids**:  `Order %in% c("Myopsida","Oegopsida")`
    - **Cetaceans**: `Order == "Artiodactyla"`
    - **Euphausiids**: `Order == "Euphausiacea"`
    - Foraminifera: n=0 for Phylum=Retaria
- Others (Phylum - Class)
    - Arthropoda
        - **Crustaceans**: `Class %in% c("Malacostraca","Ostracoda","Branchiopoda","Cephalocarida","Maxillopoda")`
        - **Sea spiders**: `Class == "Pycnogonida"`
        - skipping: Arachnida (1), Merostomata (1)
    - Mollusca
        - **Gastropods**: `Class == "Gastropoda"`
        - **Bivalves**:  `Class == "Bivalvia"`
        - **Chitons**: `Class == "Polyplacophora"`
        - Cephalopods
        - Squids
    - Chordata
        - **Tunicates**: `Class %in% c("Ascidiacea","Thaliacea")`
        - **Sharks**: `Class == "Elasmobranchii"` (actually Class=Chondrichthyes, Subclass=Elasmobranchii)
        - **Reptiles**: `Class == "Reptilia"` crocodiles, sea snakes & sea turtles
    - **Echinoderms**: `Phylum == "Echinodermata"`
    - **Sponges**: `Phylum == "Porifera"`
    - **Worms**: `Phylum == "Annelida"`
    - Cnidaria
        - **Hydrozoans**: `Class == "Hydrozoa"`
        
```{r}
library(tidyverse)

dir_dat    <- "/Volumes/Best HD/mbon_data_big/aquamaps"
spp_grouped_csv <- file.path(dir_dat, "spp_groups.csv")

spp <- suppressMessages({ read_csv(spp_grouped_csv) })

cat("\nspp$group\n")
table(spp$group)

# show AquaMap species counts by Phylum and Class ----
cat("\nspp$Phylum\n")
table(spp$Phylum)
cat("\nspp$Class\n")
table(spp$Class)
cat("\nspp$Class sorted by count\n")
table(spp$Class) %>% sort()
cat("\nspp$Class not in group\n")
spp %>% filter(is.na(group)) %>% pull(Class) %>% table() %>% sort()
```

```{r}
spp %>% 
  group_by(group) %>% 
  summarize(
    n = n() %>% scales::comma()) %>% 
knitr::kable(align = c("l","r"))
```

