---
title: "PostGIS Setup"
author: "Ben Best"
output: html_document
date: "`r format(Sys.time(), '%Y-%m-%d')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, eval=F)
```

## Data Sources

- [Marine Ecoregions of the World](http://www.marineregions.org/maps.php?album=3264&pic=64936)
- [Country by EEZ](http://www.marineregions.org/maps.php?album=3264&pic=64931)
- distance from [shore](https://www.ngdc.noaa.gov/mgg/shorelines/gshhs.html)
- [depth](http://topex.ucsd.edu/WWW_html/srtm30_plus.html)
- [sea surface temperature](https://www.ncdc.noaa.gov/oisst)
- [chlorophyll](http://oceancolor.gsfc.nasa.gov/cms/data/viirs) ([data](http://oceandata.sci.gsfc.nasa.gov/VIIRS/Mapped/Cumulative/4km/CHL_chlor_a))
- [WDPA - World Database of Protected Areas - ProtectedPlanet.net](https://protectedplanet.net/)
- [Marine Regions - downloads](http://www.marineregions.org/downloads.php)
  - [Marine Regions - webservices](http://www.marineregions.org/webservices.php)
- [Large Marine Ecosystem - Digital Data](http://www.lme.noaa.gov/index.php?option=com_content&view=category&layout=blog&id=48&Itemid=268)

```{r load}

# load libraries
suppressPackageStartupMessages({
  library(readxl)
  library(raster)
  library(rgeos)
  library(readr)
  library(tibble)
  library(dplyr)
  library(tidyr)
  library(purrr)
  library(mregions) # devtools::install_github("ropenscilabs/mregions")
  library(leaflet)
  library(htmltools)
  library(maps)
  library(mregions)
  library(ggplot2)
  library(sp)
  library(sf)
  library(rgdal)
  library(ncdf4)
  library(scales)
  library(robis)
  library(DT)
  #library(fasterize) # devtools::install_github('cohealthalliance/fasterize')
})
select = dplyr::select

# projections
crs_mol = '+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs'
crs_gcs = '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
```

## Map of OBIS by Hierarchical Hexagons

```{r map}
leaflet(wrecks_pts) %>%
  addProviderTiles('Esri.OceanBasemap') %>%  # 'Esri.OceanBasemap'/'Stamen.TonerLite'
  # see [all providers](http://leaflet-extras.github.io/leaflet-providers/preview/index.html)
  addMarkers(
    clusterOptions = T,
    popup = ~sprintf("<strong>%s</strong> (%d)<br>type: %s<br>num: %d", name, year, type, num))
```

## Map of OBIS by EEZ


## Land

```{r land}
depth_nc    = '../obis-lat-time-fig/data/GEBCO_2014_2D.nc'

depth_r = raster(depth_nc, layer = 'elevation')

system.time({
  wrecks_pts = raster::extract(depth_r, wrecks_pts, method='bilinear', sp=T)}) # 48.090 seconds
# Elevation.relative.to.sea.level
```

## Shapefiles


```bash
sudo chown -R `whoami`:admin /usr/local/share
brew link lzlib
brew install postgis
# start manually
pg_ctl -D /usr/local/var/postgres -l /usr/local/var/postgres/server.log start
# stop manually
pg_ctl -D /usr/local/var/postgres stop -s -m fast
# connect to default postgres database
psql postgres
```

Enable PostGIS

```
CREATE DATABASE shipwrecks;
-- DROP DATABASE shipwrecks;
\connect shipwrecks
-- Enable PostGIS (includes raster)
CREATE EXTENSION postgis;
-- Enable Topology
CREATE EXTENSION postgis_topology;
-- Enable PostGIS Advanced 3D
-- and other geoprocessing algorithms
-- sfcgal not available with all distributions
CREATE EXTENSION postgis_sfcgal;
-- fuzzy matching needed for Tiger
-- CREATE EXTENSION fuzzystrmatch;
-- rule based standardizer
-- CREATE EXTENSION address_standardizer;
-- example rule data set
-- CREATE EXTENSION address_standardizer_data_us;
-- Enable US Tiger Geocoder
-- CREATE EXTENSION postgis_tiger_geocoder;
```

Per [edzer/sfr: Simple features for R](https://github.com/edzer/sfr), installed dev version to get `st_write()` compatability with PostGIS vs:



uninstalling:

sudo rm -rf /Library/Frameworks/Python.framework/

- [osx - What is the safest way to removing Python framework files that are located in different place than Brew installs - Stack Overflow](http://stackoverflow.com/questions/22774529/what-is-the-safest-way-to-removing-python-framework-files-that-are-located-in-di)


So installing:

- [GDAL 2 On Mac with Homebrew](http://www.karambelkar.info/2016/10/gdal-2-on-mac-with-homebrew/)

```bash
# unlink gdal 1.x
brew unlink gdal
brew prune gdal

# remove kyngchaos.com GDAL framework install
sudo rm -rf /Library/Frameworks/GDAL.framework/
brew prune

# tap into osgeo4mac
brew tap osgeo/osgeo4mac && brew tap --repair

# install gdal2
#brew install gdal2 --with-armadillo \
#  --with-complete --with-libkml --with-unsupported
# Error: undefined method `rebuild' for #<BottleSpecification:0x007fa31d8a1bd8>
brew install gdal2

# link gdal2
brew link --force gdal2

# verify version in new terminal
gdal-config --version
# 2.1.1
gdal-config --libs
# -L/usr/local/Cellar/gdal2/2.1.1/lib -lgdal
gdal-config --cflags
# -I/usr/local/Cellar/gdal2/2.1.1/include
```

- reinstall rgdal

```R
remove.packages('rgdal')
install.packages('rgdal')
library(rgdal) # test
remove.packages('rgeos')
install.packages('rgeos')
```

- restart R session

```bash
brew info gdal2
```

```R
library(devtools)
Sys.setenv(GDAL_CONFIG='/usr/local/opt/gdal2/bin/gdal-config')
install_github('edzer/sfr')
```

screw it:

```R
install.packages("sf")
```

# load shp into db

[shp2pgsql: Using the ESRI Shapefile Loader](http://postgis.net/docs/using_postgis_dbmanagement.html#shp2pgsql_usage)

```bash
ogrinfo
cd /Users/bbest/github/shipwrecks
#shp2pgsql -I -s <SRID> <PATH/TO/SHAPEFILE> <SCHEMA>.<DBTABLE> | psql -U postgres -d <DBNAME>
# psql shipwrecks; DROP TABLE
shp2pgsql -i -D -d -s 4269 -I ./data/shipwrecks_pts.shp wrecks | psql -d shipwrecks
shp2pgsql -i -D -d -s 4269 -I ./data/World_EEZ_v9_20161021/eez.shp eez | psql -d shipwrecks
shp2pgsql -i -D -d -s 4269 -I ./data/WDPA_Dec2016-shapefile/WDPA_Dec2016-shapefile-polygons.shp wdpa | psql -d shipwrecks
```

```bash
brew install qscintilla2
brew install qgis2
```


```{r shapefiles}
library(sf)
library(RPostgreSQL)

shapes = list(
  wrecks = './data/shipwrecks_pts.shp',
  wdpa   = './data/WDPA_Dec2016-shapefile/WDPA_Dec2016-shapefile-polygons.shp', # 213,328 features
  eez    = './data/World_EEZ_v9_20161021/eez.shp',
  lme    = './data/LME66/LME66.shp',
  iho    = './data/IHO/World_Seas.shp')

shp = shapes[[2]]
obj = names(shapes[2])

db = 'shipwrecks'

con = dbConnect(dbDriver("PostgreSQL"), dbname=db, host='localhost')


ply = st_read(shp)
st_layers("PG:dbname=shipwrecks")
st_read("PG:dbname=shipwrecks")
eez = st_read_db(con, 'eez')
plot(eez)
st_write(ply, sprintf('PG:dbname=%s', db), obj, layer_options = 'OVERWRITE=true')


#select from wrecks, eez where st_intersects(wrecks.geom, eez.geom) and num in (29,30,31);

# test if column is unique
#select mrgid from eez group by mrgid having count(*) > 1;



system.time({
  res = dbGetQuery(
    con,
    "SELECT num, geoname FROM wrecks, eez WHERE ST_INTERSECTS(wrecks.geom, eez.geom);")
})
#  user  system elapsed
# 0.018   0.002   7.547

eez    = st_read(shapes[['eez']])
wrecks = st_read(shapes[['wrecks']])

system.time({
  res2 = wrecks %>%
    st_intersects(eez) %>%
    select(num, geoname)
})


ply = rgdal::readOGR(dirname(shp), tools::file_path_sans_ext(basename(shp)))

wrecks_df

pts_ply = sp::over(wrecks_pts[28:31,], ply)
pts_ply

```

```{r raster}
# raster::extract()
dir_asc = './data/BioOracle'

rasters = stack(list.files(dir_asc, '\\.asc$', full.names=T))
crs(rasters) = crs_gcs

for (asc in list.files(dir_asc, '\\.asc$', full.names=T)){ # asc = list.files(dir_asc, '\\.asc$', full.names=T)[1]


  r = rgdal::readGDAL(asc)

}
```

