---
title: "PostGIS Setup"
author: "Ben Best"
output: html_document
date: "`r format(Sys.time(), '%Y-%m-%d')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, eval=F)
```

## Technical Links

- [ecohealthalliance/fasterize: Fast vector <--> raster conversion for modern R spatial data 🚀🌏▦](https://github.com/ecohealthalliance/fasterize)

## Install Apps

Using PostGIS since big features.

Installed:

- [Postgres.app](http://postgresapp.com/) – the easiest way to get started with PostgreSQL on the Mac

- [pgAdmin](https://www.postgresql.org/ftp/pgadmin3/pgadmin4/v1.1/macos/) GUI

## Brew & Start PostGIS

```bash
sudo chown -R `whoami`:admin /usr/local/share
brew link lzlib
brew install postgis
# start manually
pg_ctl -D /usr/local/var/postgres -l /usr/local/var/postgres/server.log start
# stop manually
pg_ctl -D /usr/local/var/postgres stop -s -m fast
# connect to default postgres database
psql postgres
```

## Create Database `mbon`

### Rename Existing

For now, using existing data loaded into shipwrecks database:

```bash
psql postgres
```

```sql
ALTER DATABASE shipwrecks RENAME TO mbon;
```

```bash
\q
```

### Create New

Or, to create fresh:

```sql
CREATE DATABASE mbon;
-- DROP DATABASE mbon;
\cnnect mbon
-- Enable PostGIS (includes raster)
CREATE EXTENSION postgis;
-- Enable Topology
CREATE EXTENSION postgis_topology;
-- Enable PostGIS Advanced 3D
-- and other geoprocessing algorithms
-- sfcgal not available with all distributions
CREATE EXTENSION postgis_sfcgal;
-- fuzzy matching needed for Tiger
-- CREATE EXTENSION fuzzystrmatch;
-- rule based standardizer
-- CREATE EXTENSION address_standardizer;
-- example rule data set
-- CREATE EXTENSION address_standardizer_data_us;
-- Enable US Tiger Geocoder
-- CREATE EXTENSION postgis_tiger_geocoder;
```

Per [edzer/sfr: Simple features for R](https://github.com/edzer/sfr), installed dev version to get `st_write()` compatability with PostGIS vs:

```{r}
library(sf)
library(RPostgreSQL)

shapes <- list(
  wrecks = './data/shipwrecks_pts.shp',
  wdpa   = './data/WDPA_Dec2016-shapefile/WDPA_Dec2016-shapefile-polygons.shp', # 213,328 features
  eez    = './data/World_EEZ_v9_20161021/eez.shp',
  lme    = './data/LME66/LME66.shp',
  iho    = './data/IHO/World_Seas.shp')

shp <- shapes[[2]]
obj <- names(shapes[2])

db <- 'mbon'
con <- dbConnect(dbDriver('PostgreSQL'), dbname=db, host='localhost')

nc = st_read(system.file("shape/nc.shp", package="sf"))
st_write(nc, "nc.shp")
st_write(nc, sprintf('PG:dbname=%s', db), 'nc', layer_options = 'OVERWRITE=true')
```

```
Error in guess_driver(dsn) :
  guess_driver: PostgreSQL not available in supported drivers, see `st_drivers()'
```

uninstalling:

sudo rm -rf /Library/Frameworks/Python.framework/

- [osx - What is the safest way to removing Python framework files that are located in different place than Brew installs - Stack Overflow](http://stackoverflow.com/questions/22774529/what-is-the-safest-way-to-removing-python-framework-files-that-are-located-in-di)


So installing:

- [GDAL 2 On Mac with Homebrew](http://www.karambelkar.info/2016/10/gdal-2-on-mac-with-homebrew/)

```bash
# unlink gdal 1.x
brew unlink gdal
brew prune gdal

# remove kyngchaos.com GDAL framework install
sudo rm -rf /Library/Frameworks/GDAL.framework/
brew prune

# tap into osgeo4mac
brew tap osgeo/osgeo4mac && brew tap --repair

# install gdal2
#brew install gdal2 --with-armadillo \
#  --with-complete --with-libkml --with-unsupported
# Error: undefined method `rebuild' for #<BottleSpecification:0x007fa31d8a1bd8>
brew install gdal2

# link gdal2
brew link --force gdal2

# verify version in new terminal
gdal-config --version
# 2.1.1
gdal-config --libs
# -L/usr/local/Cellar/gdal2/2.1.1/lib -lgdal
gdal-config --cflags
# -I/usr/local/Cellar/gdal2/2.1.1/include
```

- reinstall rgdal

```R
remove.packages('rgdal')
install.packages('rgdal')
library(rgdal) # test
remove.packages('rgeos')
install.packages('rgeos')
```

- restart R session

```bash
brew info gdal2
```

```R
library(devtools)
Sys.setenv(GDAL_CONFIG='/usr/local/opt/gdal2/bin/gdal-config')
install_github('edzer/sfr')
```

screw it:

```R
install.packages("sf")
```

# load csv of points into db

Using [CSV Driver | GDAL.org](http://www.gdal.org/drv_csv.html):

``
<OGRVRTDataSource>
    <OGRVRTLayer name="test">
        <SrcDataSource>test.csv</SrcDataSource>
        <GeometryType>wkbPoint</GeometryType>
        <LayerSRS>WGS84</LayerSRS>
        <GeometryField encoding="PointFromColumns" x="Longitude" y="Latitude"/>
    </OGRVRTLayer>
</OGRVRTDataSource>

`ba``sh
csocc="/Users/bbest/github/obis-lat-time-fig/data/occurrences"
csv="$occ.csv"
vrt="$occ.vrt"

ad $csv
og
ogr2ogr --version # GDAL 1.11.3, released 2015/09/16

cd /Users/bbest/github/obis-lat-time-fig/data
vi occurrences.vrt
```

```xml
<OGRVRTDataSource>
    <OGRVRTLayer name="occurrences">
        <SrcDataSource relativeToVRT="1">occurrences.csv</SrcDataSource>
        <GeometryType>wkbPoint</GeometryType>
        <LayerSRS>WGS84</LayerSRS>
        <GeometryField encoding="PointFromColumns" x="decimalLongitude" y="decimalLatitude"/>
    </OGRVRTLayer>
</OGRVRTDataSource>
```


```bash
#ogrinfo -ro -al $csv X_POSSIBLE_NAMES=decimalLongitude Y_POSSIBLE_NAMES=decimalLatitude -oo KEEP_GEOM_COLUMNS=NO
rinfo -ro -al $c$vrt `,`

`` load EEZ, WDPA shp into db

[shp2pgsql: Using the ESRI Shapefile Loader](http://postgis.net/docs/using_postgis_dbmanagement.html#shp2pgsql_usage)

```bash
ogrinfo
cd /Users/bbest/github/mbon
#shp2pgsql -I -s <SRID> <PATH/TO/SHAPEFILE> <SCHEMA>.<DBTABLE> | psql -U postgres -d <DBNAME>
# psql mbon; DROP TABLE
shp2pgsql -i -D -d -s 4269 -I ./data/World_EEZ_v9_20161021/eez.shp eez | psql -d mbon
shp2pgsql -i -D -d -s 4269 -I ./data/WDPA_Dec2016-shapefile/WDPA_Dec2016-shapefile-polygons.shp wdpa | psql -d shipwrecks
```

## Try `sf::st_write`

```bash
brew install qscintilla2
brew install qgis2
```

```{r}
ply = st_read(shp)
st_layers("PG:dbname=shipwrecks")
st_read("PG:dbname=shipwrecks")
eez = st_read_db(con, 'eez')
plot(eez)
st_write(ply, sprintf('PG:dbname=%s', db), obj, layer_options = 'OVERWRITE=true')
```

#select from wrecks, eez where st_intersects(wrecks.geom, eez.geom) and num in (29,30,31);

# test if column is unique
#select mrgid from eez group by mrgid having count(*) > 1;

system.time({
  res = dbGetQuery(
    con,
    "SELECT num, geoname FROM wrecks, eez WHERE ST_INTERSECTS(wrecks.geom, eez.geom);")
})
#  user  system elapsed
# 0.018   0.002   7.547

eez    = st_read(shapes[['eez']])
wrecks = st_read(shapes[['wrecks']])

system.time({
  res2 = wrecks %>%
    st_intersects(eez) %>%
    select(num, geoname)
})


ply = rgdal::readOGR(dirname(shp), tools::file_path_sans_ext(basename(shp)))

wrecks_df

pts_ply = sp::over(wrecks_pts[28:31,], ply)
pts_ply

```

```{r raster}
# raster::extract()
dir_asc = './data/BioOracle'

rasters = stack(list.files(dir_asc, '\\.asc$', full.names=T))
crs(rasters) = crs_gcs

for (asc in list.files(dir_asc, '\\.asc$', full.names=T)){ # asc = list.files(dir_asc, '\\.asc$', full.names=T)[1]


  r = rgdal::readGDAL(asc)

}
```

