---
title: "ROI EBSA"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    code_folding: hide
    df_print: paged
---

## Overview

Workflow:

- select region of interest (ROI), eg CMAR EBSA
- get obis
- intersect w/ other subregions:
    - EEZ, MPAs, EBSAs,...
    - st_join obis id
- subsequent analytical aspects
    - TODO: in/out MPA. simple index?
      - proportion of obs in/out
      - filter by RedList, eg more in vs out: ratio for spp > LC (least concern)
          - drill down by # years obs & recent? **advise on future monitoring**
          - mgt msg: overlay obs grid w/ hi chl 
    - compare w/ hi chl / upwelling areas / unique seascapes
    - TODO: show ecosystem/habitats + cumulative impacts

- EEZ
    - + Disputed Territory
    - + Joint regime (EEZ)
- Territorial Sea
- Marine Protected Area (MPA)
- IHO Sea Area
- FAO Subareas
- Marine Subregion
- Marine Ecoregion of the World (MEOW)


```{r setup, warning=F, message=F}
knitr::opts_chunk$set(echo=T, message=F, warning=F)
library(tidyverse)
library(sf)
library(geosphere)
library(mapview)
library(robis)  # devtools::install_github('iobis/robis')
library(robis2) # devtools::install_github('iobis/robis2')
library(rmapshaper)
library(geojsonio)
library(scales)
library(DT)
library(leaflet)
library(htmltools)
library(htmltools)

if (basename(getwd()) != 'technical') setwd('technical')

roi              = 'cmar'
roi_url          = 'https://chm.cbd.int/api/v2013/documents/FB73B0D1-64FB-367F-405A-51AAA4C060F7/attachments/ETTP_8_EBSA.geojson'
#roi              = 'clipperton'
#roi_url          = 'https://chm.cbd.int/api/v2013/documents/494D9489-5D08-524D-84AE-AD6817304655/attachments/ETTP_2_EBSA.geojson'
roi_geo          = sprintf('cache/%s.geojson', roi)
occ_csv          = sprintf('cache/%s_occ.csv', roi)
occ_cklist_csv   = sprintf('cache/%s_occ_checklist.csv', roi)
occ_txt          = sprintf('cache/%s_occ_nwkt.txt', roi)
occ_geo          = sprintf('cache/%s_occ.geojson', roi)
eez_rdata        = 'cache/eez.rdata'
eez_geo          = sprintf('cache/%s_eez.geojson', roi)
wdpa_geo         = sprintf('cache/%s_wdpa.geojson', roi)
wdpa_eez_geo     = sprintf('cache/%s_wdpa_eez.geojson', roi)
occ_eez_wdpa_csv = sprintf('cache/%s_occ_eez_wdpa.csv', roi)

#library(mapedit)
#editMap()
```

## Fetch EBSA

Fetch EBSA polygon from [Ecologically or Biologically Significant Marine Areas (EBSAs)](https://www.cbd.int/ebsa/) -> 
[Record for Corredor Marino del Pacífico Oriental tropical](https://chm.cbd.int/database/record?documentID=204045).

```{r fetch roi poly}
if (!file.exists(roi_geo)){
  roi_sf = read_sf(roi_url)
  write_sf(roi_sf, roi_geo)  
}
roi_sf = read_sf(roi_geo)

mapview(roi_sf)
```

## Generalize EBSA for Querying OBIS

In order to query OBIS and other web services, well known text (WKT) is needed which when passed through a URL has a limited number of characters to pass. To reduce this size, buffer by 0.1 decimal degrees and simplify the polygon (by 95%).

```{r generalize roi, message=F}
roi_smp = roi_sf %>%
  st_buffer(dist=0.1) %>%
  geojson_json() %>%
  ms_simplify() %>%
  geojson_sp() %>%
  st_as_sf()

roi0_wkt = roi_sf %>%
  st_geometry() %>%
  st_as_text()

roi_wkt = roi_smp %>%
  st_geometry() %>%
  st_as_text()

mapview(roi_smp)
```
The WKT representation got reduced from `r nchar(roi0_wkt) %>% comma()` to `r nchar(roi_wkt) %>% comma()` characters.

## Fetch OBIS occurrences

Fetch OBIS records using the newer [iobis/robis2](https://github.com/iobis/robis2) package that retrieves records really fast using ElasticSearch.

```{r fetch obis tbl for roi poly}
cols_drop = c(
  'acceptedNameUsage','acceptedNameUsageID','accessRights','associatedMedia','associatedReferences','associatedSequences','associatedTaxa',
  'basisOfRecord','behavior','bibliographicCitation',
  'collectionID','continent','coordinatePrecision','coordinateUncertaintyInMeters','countryCode','county',
  'dataGeneralizations','datasetID','dateIdentified','division','dynamicProperties',
  'establishmentMeans','eventID','eventRemarks','eventTime',
  'fieldNotes','fieldNumber','footprintSRS','footprintWKT','forma',
  'geodeticDatum',
  'habitat','higherClassification','higherGeography','higherGeographyID',
  'identificationID','identificationQualifier','identificationReferences','identificationRemarks','identifiedBy','individualCount',
  'individualID','informationWithheld','infraclass','infrakingdom','infraorder','infraphylum','institutionID','island','islandGroup',
  'language','lifeStage','locality','locationAccordingTo','locationID','locationRemarks',
  'materialSampleID','maximumDepthInMeters','minimumDepthInMeters','modified','municipality',
  'nameAccordingTo','nameAccordingToID','namePublishedIn','namePublishedInID',
  'occurrenceID','occurrenceRemarks','occurrenceStatus','originalNameUsage','originalNameUsageID','otherCatalogNumbers','ownerInstitutionCode',
  'parvorder',
  'section',
  'recordNumber','recordedBy','references','reproductiveCondition','rights','rightsHolder',
  'scientificNameAuthorship','sex','source',
  'stateProvince','subclass','subdivision','subfamily','subforma','subgenus','subkingdom','suborder',
  'subphylum','subsection','subspecies','subterclass','subtribe','subvariety',
  'superclass','superfamily','superorder','supertribe',
  'taxonConceptID','taxonID','taxonRank','taxonRemarks','taxonomicStatus',
  'tribe','type','typeStatus',
  'variety','vernacularName',
  'waterBody',
  'phylum','class','order','family','genus','species') # covered by checklist

if (any(!file.exists(occ_txt), !file.exists(occ_csv))){
  occ_tbl = robis2::occurrence(geometry=roi_wkt) %>% 
    as_tibble()
  occ_tbl = occ_tbl %>%
    select_(.dots = setdiff(names(occ_tbl), cols_drop))
  
  write_lines(nrow(occ_tbl), occ_txt)
  write_csv(occ_tbl, occ_csv)
}
occ_tbl  = read_csv(occ_csv)
occ_nwkt = read_lines(occ_txt) %>% as.integer()

occ_tbl %>%
  head()
```

Above are the first 6 of `r occ_nwkt %>% comma()` records from the generalized WKT.

## Fetch Taxonomic Information with IUCN Redlist

```{r get obis checklist for roi}
if (!file.exists(occ_cklist_csv)){
  occ_cklist = robis::checklist(geometry=roi_wkt) # 1.6 min  
  
  # get unique taxonomic name (tname), prioritizing beyond with first non-NA of: valid, worms_id, redlist, status, phylum,...
  occ_cklist = occ_cklist %>%
    mutate(
      invalid = valid_id != id) %>%
    arrange(tname,invalid,worms_id,redlist,status,phylum,class,order,family,genus,species) %>%
    group_by(tname) %>%
    summarize(
      worms_id     = first(worms_id),
      checklist_id = first(id),
      valid_id     = first(parent_id),
      rank_name    = first(rank_name),
      tauthor      = first(tauthor),
      phylum       = first(phylum),
      class        = first(class),
      order        = first(order),
      family       = first(family),
      genus        = first(genus),
      species      = first(species),
      redlist      = first(redlist),
      status       = first(status),
      wktpoly_sum_records  = sum(records),
      wktpoly_sum_datasets = sum(datasets))
  
  write_csv(occ_cklist, occ_cklist_csv)
}
occ_cklist = read_csv(occ_cklist_csv)
```

```{r occ_cklist debug, eval=F}
occ_head = head(occ_tbl)
View(occ_head)
table(occ_tbl$taxonomicgroup, useNA='ifany') # 0 NAs
# occ_tbl$taxonomicgroup YES!

sum(duplicated(occ_cklist$id)) # 0
sum(duplicated(occ_cklist$worms_id)) # 154
sum(duplicated(occ_cklist$tname[duplicated(occ_cklist$tname)])) # 0

occ_cklist_worms_dupes = occ_cklist %>% 
  filter(
    worms_id %in% occ_cklist$worms_id[duplicated(occ_cklist$worms_id)] |
    tname %in% occ_cklist$tname[duplicated(occ_cklist$tname)]) %>%
  arrange(worms_id)
View(occ_cklist_worms_dupes)
```

## Convert OBIS occcurrences to points and filter

Next, convert the table having decimalLongitude and decimalLatitude into points and filter out points that were in the generalized WKT but not the original ROI polygon.

```{r convert obis tbl to pts within roi poly, message=F}
if (!file.exists(occ_geo)){
  occ_tbl = read_csv(occ_csv)

  # turn obis into sf
  occ_sf = st_sf(
    occ_tbl, 
    geometry = st_sfc(
      with(
        occ_tbl, 
        map2(decimalLongitude, decimalLatitude, function(x, y) st_point(c(x, y)))))) %>%
    st_set_crs(4326)
  
  # rm obis outside original roi: 137,277 -> 64,342
  occ_sf = occ_sf %>%
    slice(st_intersects(roi_sf, occ_sf)[[1]])

  # write filtered csv
  occ_sf %>%
    st_set_geometry(NULL) %>%
    write_csv(occ_csv)
  
  # write geo with just id to save disk space, for later rejoining with occ_tbl
  if (sum(duplicated(occ_tbl$id)) > 0) stop('Whoah! Expecting unique id in OBIS occ_tbl.')
  occ_sf %>% 
    select(id) %>%
    write_sf(occ_geo, delete_dsn=T)
}
occ_tbl = read_csv(occ_csv)
occ_sf  = read_sf(occ_geo) %>%
  left_join(
    occ_tbl, by='id')

leaflet(occ_sf) %>% 
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(
    data = roi_sf) %>%
  addMarkers(
    clusterOptions = markerClusterOptions())
```

Above are the OBIS occurrence points, filtered to the original ROI polygon (`r occ_nwkt %>% comma()` to `r nrow(occ_sf) %>% comma()` records). Click on the marker cluster to zoom and explode into finer clusters and eventually individual points.

## Fetch EEZs within ROI

Get the exclusive economic zones (EEZs) within the ROI region of interest to get at areas of political interest and management.

If we already knew the country EEZs that overlapped with the ROI region, we could use the unevaluated code below that Eduardo Klein Sala shared originally from Pieter Provost using the [mregions](https://github.com/ropenscilabs/mregions) package. 

```{r mregions from Pieter Provoost via Eduardo Klein Sala, eval=F}
library(mregions)

eez_names = mr_names('MarineRegions:eez') # View(n)
eez_id    = mr_names_search(eez_names, 'Belgian')$id[1] 
eez_json  = mr_features_get('MarineRegions:eez', eez_id, format='json') 

eez_sf = eez_json %>% 
  as.json() %>% 
  geojson_sp() %>% 
  st_as_sf()
mapview(eez_sf)
```

However am attempting to generalize this analysis so that ROI could be swapped for any region of interest. In that case we need to fetch the entire EEZ dataset (takes ~ 2 min by hitting the MarineRegions.org WFS server at VLIZ.be) and subset based on intersecting EEZs. If the `mr_records_by_type()` or other function in the mregions package were complete with the bounding box fields (`minLatitude`, `minLongitude`, `maxLatitude`, `maxLongitude`) we wouldn't need to do this.

```{r fetch entire EEZ via WFS}
if (!file.exists(eez_rdata)){
  eez_url = 'http://geo.vliz.be/geoserver/MarineRegions/ows?service=WFS&version=1.0.0&request=GetFeature&typeName=MarineRegions:eez&maxFeatures=300'
  eez_all_sf = read_sf(eez_url)       # 2.2 min
  saveRDS(eez_all_sf, file=eez_rdata) # store compressed
}
eez_all_sf = readRDS(eez_rdata)
```

```{r eez in roi, message=F}
bbox_ply = function(bb){
  st_sfc(st_polygon(list(matrix(c(
    bb[['xmin']], bb[['ymin']],
    bb[['xmin']], bb[['ymax']],
    bb[['xmax']], bb[['ymax']],
    bb[['xmax']], bb[['ymin']],
    bb[['xmin']], bb[['ymin']]), ncol=2, byrow=T)))) %>%
    st_set_crs(attr(bb,'crs'))
}

if (!file.exists(eez_geo)){

  # filter by eez's within roi
  eez_sf = eez_all_sf %>%
    slice(st_intersects(st_bbox(roi_sf) %>% bbox_ply(), eez_all_sf)[[1]]) %>%
    st_intersection(roi_sf %>% select())
  
  # update areas
  eez_sf = eez_sf %>%
    rename(
      area_eezall_km2 = area_km2) %>%
    mutate(
      area_eezroi_km2 = st_area(st_geometry(eez_sf)))
  
  write_sf(eez_sf, eez_geo, delete_dsn=T)
}
eez_sf = read_sf(eez_geo)

labels <- with(
  eez_sf,
  sprintf(
  "<strong>%s - %s</strong><br/>area_eezroi_km2: %s<br/>area_eezroi_km2: %s",
  sovereign1, territory1, comma(round(area_eezroi_km2,2)), comma(round(area_eezall_km2,2)))) %>% lapply(HTML)

leaflet(roi_sf) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons() %>%
  addPolygons(
    data = eez_sf,
    color = 'red', fillColor = 'red',
    label = labels,
    highlight = highlightOptions(
      color = 'yellow',
      weight = 5,
      bringToFront = T))
```

## Fetch MPAs in ROI

We can use the [Protected Planet](https://www.protectedplanet.net/) [ESRI Web Service Query](http://ec2-54-204-216-109.compute-1.amazonaws.com:6080/arcgis/rest/services/wdpa/wdpa/MapServer/1/query) to construct an [ESRI Query (Operation)](https://resources.arcgis.com/en/help/rest/apiref/query.html) on the bounding box of the ROI region. This could be limited to the [fields](http://ec2-54-204-216-109.compute-1.amazonaws.com:6080/arcgis/rest/services/wdpa/wdpa/MapServer/1) available.

```{r fetch wdpa within roi, message=F}
q_url = 'http://ec2-54-204-216-109.compute-1.amazonaws.com:6080/arcgis/rest/services/wdpa/wdpa/MapServer/1/query'
bb = st_bbox(roi_sf)

if (!file.exists(wdpa_geo)){
  res = httr::GET(q_url, query = list(
    f            = 'json',
    geometry     = sprintf('%0.1f,%0.1f,%0.1f,%0.1f', bb[['xmin']], bb[['ymin']], bb[['xmax']], bb[['ymax']]),
    geometryType = 'esriGeometryEnvelope',
    inSR         = 4326,
    outSR        = 4326,
    outFields    = '*'))
  
  wdpa_sf = httr::content(res) %>%
    read_sf() # wdpa0_sf = wdpa_sf; mapview(wdpa0_sf)
  
  if (nrow(wdpa_sf) == 0){
    warning('no wdpa')
    write_sf(wdpa_sf, wdpa_geo, delete_dsn=T)
  } else {
    
    # filter by eez's within roi
    wdpa_sf = wdpa_sf %>%
      slice(st_intersects(roi_sf, wdpa_sf)[[1]])
    wdpa_sf = wdpa_sf %>%
      mutate(
        area_wdpaall_km2 = st_area(st_geometry(wdpa_sf)))
    wdpa_sf = wdpa_sf %>%
      st_intersection(roi_sf %>% select())
    wdpa_sf = wdpa_sf %>%
      mutate(
        area_wdparoi_km2 = st_area(st_geometry(wdpa_sf)))
  
    write_sf(wdpa_sf, wdpa_geo, delete_dsn=T)
  }
}
wdpa_sf = read_sf(wdpa_geo)

if (nrow(wdpa_sf) == 0){
  warning('no wdpa')
} else {  
  labels <- with(
    wdpa_sf,
    sprintf(
    "<strong>%s</strong><br/>DESIG_ENG: %s<br/>DESIG_TYPE: %s<br/>IUCN CAT: %s<br/>GIS_M_AREA: %s",
    NAME, DESIG_ENG, DESIG_TYPE, IUCN_CAT, comma(round(GIS_M_AREA,2)))) %>% lapply(HTML)
  
  leaflet(roi_sf) %>%
    addProviderTiles(providers$Stamen.TonerLite) %>%
    addPolygons() %>%
    addPolygons(
      data = wdpa_sf,
      weight = 1,
      color = 'green', fillColor = 'green',
      label = labels,
      highlight = highlightOptions(
        color = 'yellow',
        weight = 5,
        bringToFront = T))
}
```

## MPAs in EEZs

```{r}
if (!file.exists(wdpa_eez_geo)){
  wdpa_eez_sf = wdpa_sf %>%
    st_intersection(eez_sf)
  
  write_sf(wdpa_eez_sf, wdpa_eez_geo, delete_dsn=T)
}
wdpa_eez_sf = read_sf(wdpa_eez_geo)

labels = with(
  wdpa_eez_sf,
  sprintf(
  "<strong>%s</strong> in <strong>%s - %s</strong><br/>DESIG_ENG: %s<br/>DESIG_TYPE: %s<br/>IUCN CAT: %s<br/>GIS_M_AREA: %s",
  NAME, sovereign1, territory1, DESIG_ENG, DESIG_TYPE, IUCN_CAT, comma(round(GIS_M_AREA,2)))) %>% lapply(HTML)

wdpa_eez_sf$sovereign1 = factor(wdpa_eez_sf$sovereign1)
pal = colorFactor('Spectral', wdpa_eez_sf$sovereign1)

leaflet(roi_sf) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(
    weight = 0.5,
    opacity = 0.2) %>%
  addPolygons(
    data = wdpa_eez_sf,
    weight = 1,
    fillColor = ~pal(sovereign1), # color = ~pal(sovereign1), 
    label = labels,
    highlight = highlightOptions(
      color = 'yellow',
      weight = 5,
      bringToFront = T))
```


## Spatial Join EEZ & WDPA to OBIS occurrences

Spatially identify the EEZ and WDPA for each OBIS occurrence.

```{r obis by eez & wdpa}
if (!file.exists(occ_eez_wdpa_csv)){
  o_sf = occ_sf %>%
    select(occ_id=id) %>%
    st_join(
      eez_sf %>% 
        select(eez_gml_id=gml_id)) %>%
    st_join(
      wdpa_sf %>% 
        select(wdpa_pid = WDPA_PID)) # 4.9 min
  
  o_sf %>%
    st_set_geometry(NULL) %>%
    write_csv(occ_eez_wdpa_csv)
}
occ_eez_wdpa = read_csv(occ_eez_wdpa_csv)

occ_eez_wdpa %>%
  head()
```

## Summarize OBIS by EEZ, WDPA, IUCN

```{r}
o = occ_tbl %>% 
  left_join(
    occ_cklist, 
    by=c('scientificName'='tname')) %>%
  left_join(
    occ_eez_wdpa, 
    by=c('id'='occ_id')) %>%
  left_join(
    eez_sf %>%
      st_set_geometry(NULL), 
    by=c('eez_gml_id'='gml_id')) %>%
  left_join(
    wdpa_sf %>%
      mutate(WDPA_PID = as.integer(WDPA_PID)) %>%
      st_set_geometry(NULL), 
    by=c('wdpa_pid'='WDPA_PID'))

o %>%
  group_by(taxonomicgroup) %>%
  summarize(n = n()) %>%
  arrange(desc(n)) %>%
  datatable() %>%
    formatCurrency('n', currency='', digits=0) 

table(o$status, useNA = 'ifany')

o %>%
  #filter(status %in% c('NT','VU','EN','CR','EX')) %>% # skipping: <NA>,DD,LC
  group_by(taxonomicgroup, status) %>%
  summarize(n = n()) %>%
  left_join(
    o %>%
      group_by(taxonomicgroup) %>%
      summarize(n_taxonomicgroup = n()),
    by='taxonomicgroup') %>%
  arrange(desc(n_taxonomicgroup, n)) %>%  
  datatable() %>%
    formatCurrency(c('n','n_taxonomicgroup'), currency='', digits=0)
```



